---
title: "IBM Data Science Skills Network: R and RStudio "
output: html_notebook
---

# Lab 1:Creating Data Visualizations using ggplot

### Objective

-   Create different data visualizations using the ggplot package with inbuilt data set mtcars.

##### 

## Loading Data Set

```{r}
library(datasets)
data("mtcars")
View(mtcars)
```

##### 

### Getting more info about the data set 'mtcars'

```{r}
?mtcars
```

##### 

### Installing ggplot2

```{r}
# Installing package 'ggplot2'
install.packages("ggplot2")

# Loading ggplot package
library(ggplot2)
```

##### 

## Plotting

```{r}
ggplot(aes(x=disp,y=mpg,), data=mtcars)+
  geom_point()+ #value is represented by a point
  ggtitle("Figure 1.1: Displacement vs miles per gallon")
```

### 

### Stylizing the plots

```{r}
# creates a plot named 'p'
p <- ggplot(mtcars, aes(x=disp, y=mpg))+
  geom_point()+
  ggtitle("Figure 1.2: Displacement vs Miles per gallon")+
  labs(x="Displacement", y="Miles per gallon")

# Modification of the plot's visual theme with the theme function.
p + theme(plot.title = element_text(size=14)) # the addition component adds the modified plot back to the ggplot object named 'p
```

### Changing axis name of plots

```{r}
ggplot(aes(x=disp,y=mpg,), data=mtcars)+
  geom_point()+ #value is represented by a point
  ggtitle("Figure 1.3: Displacement vs miles per gallon")+
  labs(x="Displacement", y="Miles per gallon") #Change in axis name 

```

##### 

### Creating a box plot of the distribution of mile per gallon (mpg) for the individual engine types (vs)

```{r}
mtcars$vs <- as.numeric(mtcars$vs)
ggplot(mtcars, aes(x=vs, y=mpg))+ 
  geom_boxplot()+
  ggtitle("Figure 1.4: An example boxplot of the distribution of mile per gallon against enegine types")
       
```

The above box plot function only generated 1 box plot which includes both engine types. Why is that? and Why it may not be suitable?

-   The question asked for separate box plots for each engine type.
-   The variable 'VS' is not continuous, it is binary categorical.

To fix this, change the binary categorical data to factor or string.

Instances where data type conversion is needed:

-   Grouping: Where different box plot is needed for each group in a categorical variable. In this case a factor / string needs to represent the groupings

-   Labels: When the x-axis label is numerical but text labels are prefered.

-   Ordering: If the order of the boxes in the plot needs to be change to meet some criterion other than numerical values

### Using factors to create box plots

```{r}
# factor the values of the variable 'vs' in the dataset 'mtcars', take the result and replace the existing data in the variable 'vs' in the dataset 'mtcars'
mtcars$vs<-as.factor(mtcars$vs) 
```

```{r}
p <- ggplot(mtcars, aes(x=vs,y=mpg))+
  geom_boxplot()+
  ggtitle("Figure 1.5: An example boxplot of the distribution of mile per gallon against engine types")

p + theme(plot.margin = unit(c(1, 1, 1.5, 1), "lines"), 
          #adjust margin of plots with unit (top,right,bottom,left)
          
          plot.title.position = "plot", 
          #position = plot ensures that the title is positioned within the plot area
          
          plot.title = element_text(hjust = 0.5, size = 12), 
          #hjust is horizontal alignment
          
          plot.background = element_rect(fill = "white"))
```

This box plot seems more accurate than the previous one.

### Changing colours in a plot

```{r}
p <- ggplot(mtcars, aes(x=vs,y=mpg))+
  geom_boxplot(color = 'blue') +
  ggtitle("Figure 1.6: An example boxplot of the distribution of mile per gallon against engine types")

p + theme(plot.margin = unit(c(1, 1, 1.5, 1), "lines"), 
          #adjust margin of plots with unit (top,right,bottom,left)
          plot.title.position = "plot", 
          #position = plot ensures that the title is positioned within the plot area
          
          plot.title = element_text(hjust = 0.5, size = 12), 
          #hjust is horizontal alignment
          
          plot.background = element_rect(fill = "white"))
```

```{r}
# Assigning colur based on grouping within a boxplot

p <- ggplot(mtcars, aes(x=vs,y=mpg, fill=vs))+
  geom_boxplot() +
  ggtitle(
    "Figure 1.7: An example boxplot of the distribution of mile per gallon against engine types")+
  scale_fill_manual(values = c("#2b303b", "#aeb6bf"))


p + theme(plot.margin = unit(c(1, 1, 1.5, 1), "lines"), 
          #adjust margin of plots with unit (top,right,bottom,left)
          plot.title.position = "plot", 
          #position = plot ensures that the title is positioned within the plot area
          
          plot.title = element_text(hjust = 0.5, size = 12), 
          #hjust is horizontal alignment
          
          plot.background = element_rect(fill = "white"))

```

## Creating Histogram

```{r}
ggplot(mtcars, aes(x=wt))+
  geom_histogram(binwidth = 0.5)+
  ggtitle("Figure 1.8: A histogram of the weight of cars")
```
